Dependently Typed Multi-stage Calculus

### 2019-11-27 slides/main.tex
Add.

\begin{frame}[fragile]{What is Multi-stage Programming}
    A programming paradigm that enables
    \begin{itemize}
        \item Generation of code at runtime
        \item Evaluation of code at runtime
    \end{itemize}
    Applications
    \begin{itemize}
        \item Speeding up programs using runtime information
    \end{itemize}
    \note{
    }
\end{frame}

\begin{frame}[fragile]{Example of Multi-stage Programming in MetaOCaml}
    \begin{exampleblock}{}
        \begin{verbatim}
 # let a = < (1 + 2) >;;
 val a : int code = < (1 + 2) >
 # run a;;
 - : int = 3

 # let b = <(~a) + (~a)>
 val b : int code = < ((1 + 2) + (1 + 2)) >
 # run b
 - : int = 6
        \end{verbatim}
    \end{exampleblock}
    \note{
    }
\end{frame}

@@

\begin{frame}[fragile]{\LTP [Hanada\&Igarashi'14]}
    \begin{table}
        \begin{tabular}{ c | c | c }
            & \LTP & MetaOCaml \\[2mm]
            \hline
            Quote & \( \textbf{let } a = \TB (1 + 2) \) & \verb| let a = <(1+2)> | \\[2mm]
            Unquote & \( (\TB (\TBL a + \TBL a)) \) & \verb| <(~a)+(~a)> | \\[2mm]
            Run & \( \textbf{run } (\TB (1 + 2)) \) & \verb| run <(1+2)> | \\[2mm]
            Code type & \( \TW\ \text{Int} \) & \verb| int code |
        \end{tabular}
    \end{table}
\end{frame}

@@

\begin{frame}[fragile]{Types of \LTP}
    \begin{block}{Types}
        \begin{tabbing}
            \hspace{5mm} \( \tau \) \= ::= X \hspace{20mm} \= (type-level constants) \\
            \> \( \mid \TW\ \tau \) \> (code type) \\
        \> \( \mid \tau \to \tau \) \> (function type)
    \end{tabbing}
    \end{block}
   \note{
    }
\end{frame}

@@

\section{Extend \LTP with Dependent Types}

@@

\begin{frame}[fragile]{Solution}
    \newcommand{\Vt}{\text{Vec}\ 3}
    \newcommand{\Vpt}{\text{Vec}\ \%3}
    Erase the CSP symbol \( \% \) of an closed term.
    \begin{center}
        \infrule[{\QPercent}]{
            M \text{ is a closed term.}
        }{
            \G\vdash \% M \E M
        }
    \end{center}
    \begin{exampleblock}{}
    \begin{itemize}
        \item \( \text{Vec}\ \%3 \E \text{Vec}\ 3 \)
        \item \( \text{vadd3} : \TW (\Vt \to \Vt \to \Vt) \) \\
            \hspace{15mm} \( (\E  \TW (\Vpt \to \Vpt \to \Vpt)) \)
    \end{itemize}
    \end{exampleblock}
\end{frame}

@@

>>>>

\begin{frame}[fragile]{What is Multi-stage Programming}
    A programming paradigm that enables
    \begin{itemize}
        \item Generation of code at run-time
        \item Evaluation of code at run-time
    \end{itemize}
    Applications
    \begin{itemize}
        \item Speeding up programs using run-time information
    \end{itemize}
    \note{
    }
\end{frame}

\begin{frame}[fragile]{Example of Multi-stage Programming in MetaOCaml}
    \begin{exampleblock}{}
        \begin{verbatim}
 # let a = < (1 + 2) >     (* Make a code *)
 val a : int code = < (1 + 2) >
 # run a                   (* Run a code *)
 - : int = 3

 # let b = <(~a) + (~a)>   (* Embed codes *)
 val b : int code = < ((1 + 2) + (1 + 2)) >
 # run b
 - : int = 6
        \end{verbatim}
    \end{exampleblock}
    \note{
    }
\end{frame}

@@

\begin{frame}[fragile]{Quick Overview of \LTP [Hanada\&Igarashi'14]}
    \begin{table}
        \begin{tabular}{ c | c | c }
            & \LTP & MetaOCaml \\[2mm]
            \hline
            Quote & \( \textbf{let } a = \TB (1 + 2) \) & \verb| let a = <(1+2)> | \\[2mm]
            Unquote & \( (\TB (\TBL a + \TBL a)) \) & \verb| <(~a)+(~a)> | \\[2mm]
            Run & \( \textbf{run } (\TB (1 + 2)) \) & \verb| run <(1+2)> | \\[2mm]
            Code type & \( \TW\ \text{Int} \) & \verb| int code |
        \end{tabular}
    \end{table}
\end{frame}

@@

\begin{frame}[fragile]{Types of \LTP}
    \begin{block}{Types}
        \begin{tabbing}
            \hspace{5mm} \( \tau \) \= ::= X \hspace{20mm} \= (type-level constants) \\
            \> \( \mid \TW\ \tau \) \> (code type) \\
        \> \( \mid \tau \to \tau \) \> (function type)
    \end{tabbing}
    \end{block}

    \begin{block}{Typing Judgement}
        \begin{center}
            \( \G \V M : \tau @ n \)
        \end{center}
        * \( n \) is a stage.
    \end{block}

    \note{
    }
    \end{frame}

@@

\section{Extension \LTP with Dependent Types}

@@

\begin{frame}[fragile]{Solution}
    \newcommand{\Vt}{\text{Vec}\ 3}
    \newcommand{\Vpt}{\text{Vec}\ \%3}
    Erase the CSP symbol \( \% \) under the condition \(M \) is closed.
    \begin{center}
        \infrule[{\QPercent}]{
            M \text{ is a closed term.}
        }{
            \G\vdash \% M \E M
        }
    \end{center}
    \begin{exampleblock}{}
    \begin{itemize}
        \item \( \text{Vec}\ \%3 \E \text{Vec}\ 3 \)
        \item \( \text{vadd3} : \TW (\Vt \to \Vt \to \Vt) \) \\
            \hspace{15mm} \( (\E  \TW (\Vpt \to \Vpt \to \Vpt)) \)
    \end{itemize}
    \end{exampleblock}
\end{frame}

@@

\begin{frame}[fragile]{Why a Stage is Needed?}
    Without stage, \textbf{run} may causes a run-time error.
    \begin{exampleblock}{}
        \begin{center}
    \( \TB (\lambda x:\tau. \TBL (\textbf{run}\ \TB x)) \)
        \end{center}
    \end{exampleblock}

    With stage, the same code cannot be typed.\\
    Because the subterm \( (\Lambda\alpha.\ \TB_\alpha x) \) is ill-typed.
    \begin{exampleblock}{}
        \begin{center}
    \( \TB_\alpha (\lambda x:\tau. \TBL_\alpha ((\Lambda\alpha.\ \TB_\alpha x)\ \varepsilon)) \)
        \end{center}
    \end{exampleblock}
\end{frame}
